---
title: "Learning x Motivation Main Analysis and Figures"
author: "Calum Guinea"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)
```

# Overview

This notebook provides a complete workflow for analyzing Learning x Motivation (LxM) data including:

- Data loading and preprocessing
- Mixed model fitting and comparison
- Publication-quality visualizations (exact replications of original code)
- Statistical analyses

**To use this notebook:** keep this file and the `data/` folder together, then click "Knit". No paths to edit.

```
your-folder/
├── LxM_analysis_mainFigs.Rmd
└── data/
    ├── choice_trialByTrial.csv
    ├── summaryFull.csv
    └── allEstsByStim.csv
```

---

# Setup

## Locate the Data Folder

```{r locate-data-folder}
# Folder holding the raw CSVs, next to this .Rmd file
data_dir <- file.path("data")
required_files <- c("choice_trialByTrial.csv", "summaryFull.csv", "allEstsByStim.csv")

# Fail early with a clear message if a file is missing, rather than a
# cryptic error later in the notebook
missing_files <- required_files[!file.exists(file.path(data_dir, required_files))]
if (length(missing_files) > 0) {
  stop("Missing data file(s): ", paste(missing_files, collapse = ", "),
       "\nExpected them in: ", normalizePath(data_dir, mustWork = FALSE))
}
```

## Install Packages (if needed)

Run this chunk once if you need to install packages:

```{r install-packages, eval=FALSE}
install.packages(c("dplyr", "tidyr", "lme4", "ggplot2", "ggeffects", 
                   "broom.mixed", "ggdist", "ggpubr", "ppcor", "corrplot",
                   "reshape2", "knitr", "DT", "gridExtra"))
```

## Load Packages

```{r load-packages}
library(dplyr)
library(tidyr)
library(lme4)
library(ggplot2)
library(ggeffects)
library(broom.mixed)
library(ggdist)
library(ggpubr)
library(ppcor)
library(corrplot)
library(reshape2)
# library(knitr)
library(DT)
library(gridExtra)
```

---

# Data Loading and Preprocessing

## Define Preprocessing Function

This function loads the data files and prepares them for analysis by:

- Converting valence to binary (0 = loss, 1 = reward)
- Scaling continuous predictors (mean-centering and standardizing)
- Optionally excluding participants with poor tournament performance

```{r preprocessing-function}
load_and_preprocess_data <- function(exclude_low_performance = TRUE) {
  
  # Read data files from the data/ subfolder (built above as `data_dir`)
  tbtDf <- read.csv(file.path(data_dir, "choice_trialByTrial.csv"))
  summDf <- read.csv(file.path(data_dir, "summaryFull.csv"))
  
  # Rename ID column for consistency
  summDf <- summDf %>% rename(prolificID = ParticipantId)
  
  # Convert valence to binary factor (0 = loss, 1 = reward)
  tbtDf <- tbtDf %>% mutate(valence = ifelse(valence == -1, 0, 1))
  tbtDf$valence <- as.factor(tbtDf$valence)
  
  # Scale continuous predictors (mean-centered and standardized)
  # Learning fidelity measures
  tbtDf$postLearnRat_resc <- scale(tbtDf$postLearnRat)
  tbtDf$postTournRat_resc <- scale(tbtDf$postTournRat)
  tbtDf$postEffortRat_resc <- scale(tbtDf$postEffortRat)
  tbtDf$meanEstByStim_resc <- scale(tbtDf$meanEstByStim)
  tbtDf$finLearningEst_resc <- scale(tbtDf$finLearningEst)
  
  # Decision variables
  tbtDf$effDiscEV <- scale(tbtDf$effDiscEV)
  tbtDf$alienEV <- scale(tbtDf$alienEV)
  
  # Demographics
  tbtDf$Age_resc <- scale(tbtDf$Age)
  tbtDf$Sex <- as.factor(tbtDf$Sex)
  tbtDf$Gender <- as.factor(tbtDf$Gender)
  
  # Questionnaire scores
  tbtDf$SHAPS_resc <- scale(tbtDf$SHAPS)
  summDf$SHAPS_resc <- scale(summDf$SHAPS)
  tbtDf$AD_resc <- scale(tbtDf$AD)
  tbtDf$SW_resc <- scale(tbtDf$SW)
  tbtDf$Compul_resc <- scale(tbtDf$Compul)
  tbtDf$FAS_resc <- scale(tbtDf$FAS)
  tbtDf$AES_resc <- scale(tbtDf$AES)
  tbtDf$STAI_resc <- scale(tbtDf$STAI)
  
  summDf$FAS_resc <- scale(summDf$FAS)
  summDf$AES_resc <- scale(summDf$AES)
  summDf$STAI_resc <- scale(summDf$STAI)
  
  # Exclude participants with poor tournament performance if requested
  if (exclude_low_performance) {
    threshold <- 50  # Above chance performance
    tbtDf <- tbtDf[tbtDf$ovrperc > threshold, ]
    summDf <- summDf[summDf$ovrperc > threshold, ]
  }
  
  # Create expected value difference variable
  tbtDf <- tbtDf %>%
    mutate(evDiff = ifelse(
      outMag < 0,
      (abs(outMag) * outProb) - (abs(outMag) * 1),
      (abs(outMag) * outProb) - (abs(outMag) * 0)
    ))
  tbtDf$evDiff <- abs(tbtDf$evDiff)
  
  return(list(tbtDf = tbtDf, summDf = summDf))
}
```

## Load Data

```{r load-data}
data <- load_and_preprocess_data(exclude_low_performance = TRUE)
tbtDf <- data$tbtDf
summDf <- data$summDf
```

## Sample Characteristics

```{r sample-stats}
cat(sprintf("Sample size: %d participants\n", length(unique(tbtDf$prolificID))))
cat(sprintf("Total trials: %d\n", nrow(tbtDf)))
cat(sprintf("Mean age: %.1f (SD = %.1f)\n", 
            mean(summDf$Age, na.rm = TRUE), sd(summDf$Age, na.rm = TRUE)))
```

## Baseline Model Comparison

This section fits a baseline GLM without random effects to test whether the fixed effects 
significantly improve model fit compared to a null (intercept-only) model.

```{r baseline-glm}
# Fit baseline model with all fixed effects (no random effects)
# This model includes demographics and all decision variables
baseGlm <- glm(accepted ~ Age_resc + Gender + SES + ord + outProb + normOutMag + effPrp + valence,
               data = tbtDf,
               family = "binomial")

# Display model summary
summary(baseGlm)
```

```{r null-glm}
# Fit null model (intercept only)
# This serves as a baseline for comparison
nullGlm <- glm(accepted ~ 1,
               data = tbtDf,
               family = "binomial")

# Display null model summary
summary(nullGlm)
```

```{r glm-comparison}
# Compare baseline model to null model using likelihood ratio test
# Tests whether adding predictors significantly improves model fit
anova(baseGlm, nullGlm, test = "Chi")
```

---

# Model Fitting

## Define Model Formulas

We compare two sets of models:

1. **Learning measure models**: Compare different ways of measuring individualized learning
2. **Mental health models**: Test effects of psychiatric symptoms on effort-based decisions

```{r model-formulas}
# Set 1: Compare different individualized learning measures
learning_models <- list(
  base = accepted ~ Age_resc + Gender + outProb + normOutMag + effPrp + valence + 
    (1 + outProb + normOutMag + effPrp + valence | prolificID),
  
  post_learn = accepted ~ Age_resc + Gender + postLearnRat + normOutMag + effPrp + valence + 
    (1 + postLearnRat + normOutMag + effPrp + valence | prolificID),
  
  post_tourn = accepted ~ Age_resc + Gender + postTournRat + normOutMag + effPrp + valence + 
    (1 + postTournRat + normOutMag + effPrp + valence | prolificID),
  
  mean_estimate = accepted ~ Age_resc + Gender + meanEstByStim + normOutMag + effPrp + valence + 
    (1 + meanEstByStim + normOutMag + effPrp + valence | prolificID),
  
  final_learning = accepted ~ Age_resc + Gender + finLearningEst + normOutMag + effPrp + valence + 
    (1 + finLearningEst + normOutMag + effPrp + valence | prolificID)
)

# Set 2: Mental health predictors
mental_health_models <- list(
  base = accepted ~ Age_resc + Gender + postTournRat + normOutMag + effPrp + valence + 
    (1 + postTournRat + normOutMag + effPrp + valence | prolificID),
  
  shaps = accepted ~ Age_resc + Gender + postTournRat + normOutMag + effPrp + valence + SHAPS_resc + 
    (1 + postTournRat + normOutMag + effPrp + valence | prolificID),
  
  transdiagnostic = accepted ~ Age_resc + Gender + postTournRat + normOutMag + effPrp + valence + 
    AD + Compul + SW + (1 + postTournRat + normOutMag + effPrp + valence | prolificID),
  
  fatigue = accepted ~ Age_resc + Gender + postTournRat + normOutMag + effPrp + valence + FAS_resc + 
    (1 + postTournRat + normOutMag + effPrp + valence | prolificID)
)
```

## Model Fitting Function

This function fits all models in a set and compares them using AIC and BIC.

```{r model-fitting-function}
fit_model_set <- function(data, model_formulas, set_name = "models") {
  
  cat(sprintf("Fitting %s...\n", set_name))
  
  models <- list()
  results <- data.frame(
    model = character(),
    AIC = numeric(),
    BIC = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (i in seq_along(model_formulas)) {
    model_name <- names(model_formulas)[i]
    cat(sprintf("  Fitting model %d/%d: %s\n", i, length(model_formulas), model_name))
    
    model <- glmer(
      model_formulas[[i]], 
      data = data, 
      family = binomial(link = "logit"),
      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 2e5))
    )
    
    models[[model_name]] <- model
    results <- rbind(results, data.frame(
      model = model_name,
      AIC = AIC(model),
      BIC = BIC(model)
    ))
  }
  
  # Identify best model
  best_model_name <- results$model[which.min(results$AIC)]
  cat(sprintf("\nBest model by AIC: %s\n", best_model_name))
  
  return(list(
    models = models,
    results = results,
    best_model = models[[best_model_name]]
  ))
}
```

## Fit Learning Models

```{r fit-learning-models}
learning_fits <- fit_model_set(tbtDf, learning_models, "learning measure models")
```

### Model Comparison - Learning Models

```{r learning-comparison-table}
DT::datatable(learning_fits$results, 
          caption = "Learning Models Comparison",
          options = list(pageLength = 10))
```

```{r learning-comparison-plots, fig.height=4}
# AIC comparison
p_aic <- ggplot(learning_fits$results, aes(x = reorder(model, AIC), y = AIC)) +
  geom_col(fill = "steelblue") +
  labs(x = "Model", y = "AIC") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# BIC comparison
p_bic <- ggplot(learning_fits$results, aes(x = reorder(model, BIC), y = BIC)) +
  geom_col(fill = "darkgreen") +
  labs(x = "Model", y = "BIC") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

gridExtra::grid.arrange(p_aic, p_bic, ncol = 2)
```

### Best Learning Model Summary

```{r learning-model-summary}
summary(learning_fits$best_model)
```

## Fit Mental Health Models

```{r fit-mh-models}
mh_fits <- fit_model_set(tbtDf, mental_health_models, "mental health models")
```

### Model Comparison - Mental Health Models

```{r mh-comparison-table}
DT::datatable(mh_fits$results, 
          caption = "Mental Health Models Comparison",
          options = list(pageLength = 10))
```

```{r mh-comparison-plots, fig.height=4}
# AIC comparison
p_aic_mh <- ggplot(mh_fits$results, aes(x = reorder(model, AIC), y = AIC)) +
  geom_col(fill = "steelblue") +
  labs(x = "Model", y = "AIC") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# BIC comparison
p_bic_mh <- ggplot(mh_fits$results, aes(x = reorder(model, BIC), y = BIC)) +
  geom_col(fill = "darkgreen") +
  labs(x = "Model", y = "BIC") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

gridExtra::grid.arrange(p_aic_mh, p_bic_mh, ncol = 2)
```

### Best Mental Health Model Summary

```{r mh-model-summary}
summary(mh_fits$best_model)
```

---

# Visualizations - Model Effects

## Fixed Effects

Fixed effects show the average impact of each predictor across all participants.

```{r fixed-effects-plot, fig.width=9, fig.height=6}
# Extract fixed effects from best learning model
fixed_effects <- tidy(learning_fits$best_model, effects = "fixed", conf.int = TRUE, std.err = TRUE)

# Convert term to factor and explicitly set the order to match the original data
fixed_effects$term <- factor(fixed_effects$term, levels = fixed_effects$term)

# Create a vibrant color palette with unique colors for each bar
vibrant_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "gold", "#A65628", "#F781BF", "#1B9E77")

# Create the bar plot with ordered factors and color updates
winMod_fixEffBar <- ggplot(fixed_effects, aes(x = term, y = estimate, fill = term)) +
  geom_bar(stat = "identity", color = "black", width = 0.5, linewidth = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "black", linewidth = 0.8) +
  scale_fill_manual(values = vibrant_colors) +
  scale_x_discrete(labels = c("(Intercept)","Age","Gender Non-binary","Gender Not specified","Gender Woman","Post-tournament rating","Outcome magnitude","Effort level","Valence (effect of reward)")) +
  scale_y_continuous(seq(-15,15,by = 2)) +
  labs(
    x = "Fixed Effects",
    y = "Beta estimate") +
  lims(
    y = c(-15,15)
  ) +
  theme_minimal(base_size = 18) +
  theme(
    axis.text.x = element_text(size = 18, angle = 45, hjust = 1),
    legend.position = "none"
  )
print(winMod_fixEffBar)
```

## Random Effects

Random effects capture individual differences in how predictors affect choice behavior.

```{r random-effects-extraction}
# Extract random effects
random_effects <- ranef(learning_fits$best_model)$prolificID

# Assign random effects for each predictor
outProbEstRfx <- setNames(as.data.frame(random_effects$postTournRat), "outProbRfx")
effLvlRfx <- setNames(as.data.frame(random_effects$effPrp), "effRfx")
outMagRfx <- setNames(as.data.frame(random_effects$normOutMag), "outMagRfx")
valRfx <- setNames(as.data.frame(random_effects$valence1), "valRfx")
```

### Distribution of Random Slopes

```{r random-slopes-post-tournament, fig.width=4, fig.height=6}
# Post-tournament rating slopes
bin_count <- 20
x_min <- min(outProbEstRfx)
x_max <- max(outProbEstRfx)
bin_width <- (x_max - x_min) / bin_count

hist_postTourn <- ggplot(outProbEstRfx, aes(x = random_effects$postTournRat)) + 
  geom_histogram(binwidth = bin_width, fill = "#66C2A5", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.3, x_max*1.3)) +
  labs(
    x = "Post-tourn. rat. slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)
print(hist_postTourn)
```

```{r random-slopes-effort, fig.width=4, fig.height=6}
# Effort level slopes
x_min <- min(effLvlRfx)
x_max <- max(effLvlRfx)
bin_width <- (x_max - x_min) / bin_count

hist_effort <- ggplot(effLvlRfx, aes(x = random_effects$effPrp)) + 
  geom_histogram(binwidth = bin_width, fill = "#FC8D62", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.3, x_max*1.3)) +
  labs(
    x = "Effort level slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)
print(hist_effort)
```

```{r random-slopes-magnitude, fig.width=4, fig.height=6}
# Outcome magnitude slopes
x_min <- min(outMagRfx)
x_max <- max(outMagRfx)
bin_width <- (x_max - x_min) / bin_count

hist_outMag <- ggplot(outMagRfx, aes(x = random_effects$normOutMag)) + 
  geom_histogram(binwidth = bin_width, fill = "#FEE08B", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.3, x_max*1.3)) +
  labs(
    x = "Outc. magn. slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)
print(hist_outMag)
```

---

# Mental Health Effects

## SHAPS Main Effect

```{r shaps-main-effect, fig.width=8, fig.height=5}
# Check if SHAPS is in the model
if ("SHAPS_resc" %in% names(coef(mh_fits$best_model)$prolificID)) {
  
  ggpredict(mh_fits$best_model, terms = "SHAPS_resc [all]") %>%
    plot() +
    labs(x = "SHAPS score (z-scored)", y = "P(accept)",
         title = "Effect of Anhedonia on Effort Acceptance") +
    theme_minimal(base_size = 14)
  
} else {
  cat("SHAPS not included in best mental health model\n")
}
```

## Mental Health Model Visualizations

### Fixed Effects Bar Plot

Visualizes the fixed effect estimates from the best mental health model.

```{r mh-fixed-effects-extraction}
# Extract fixed and random effects from the best mental health model
fixed_effects_mh <- fixef(mh_fits$best_model)
random_effects_mh <- ranef(mh_fits$best_model)$prolificID

# Create data frames for each random effect slope
# These capture individual differences in how predictors affect decisions
outProbEstRfx <- setNames(as.data.frame(random_effects_mh$postTournRat), "outProbRfx")
effLvlRfx <- setNames(as.data.frame(random_effects_mh$effPrp), "effRfx")
outMagRfx <- setNames(as.data.frame(random_effects_mh$normOutMag), "outMagRfx")
valRfx <- setNames(as.data.frame(random_effects_mh$valence1), "valRfx")
```

```{r mh-fixed-effects-plot, fig.width=9, fig.height=6}
# Extract fixed effects with confidence intervals for plotting
fixed_effects_mh_tidy <- tidy(mh_fits$best_model, effects = "fixed", conf.int = TRUE, std.err = TRUE)

# Convert term to factor to maintain order in plot
fixed_effects_mh_tidy$term <- factor(fixed_effects_mh_tidy$term, levels = fixed_effects_mh_tidy$term)

# Create vibrant color palette - one color per fixed effect
vibrant_colors <- c("#D7191C", "#FDAE61", "#ABDDA4", "#2B83BA", "#F46D43", 
                    "#74ADD1", "#FEE08B", "#66C2A5", "#A6D854", "#E78AC3")

# Create bar plot showing fixed effect estimates with error bars
mhWinMod_fixEffBar <- ggplot(fixed_effects_mh_tidy, aes(x = term, y = estimate, fill = term)) +
  geom_bar(stat = "identity", color = "black", width = 0.5, linewidth = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "black", linewidth = 0.8) +
  scale_fill_manual(values = vibrant_colors) +
  scale_x_discrete(labels = c("(Intercept)", "Age", "Gender Non-binary", "Gender Not specified", "Gender Woman",
                              "Post-tournament rating", "Outcome magnitude", "Effort level",
                              "Valence (effect of reward)", "SHAPS Sum score")) +
  scale_y_continuous(seq(-15, 15, by = 2)) +
  labs(
    x = "Fixed Effects",
    y = "Beta estimate") +
  lims(y = c(-15, 15)) +
  theme_minimal(base_size = 18) +
  theme(
    axis.text.x = element_text(size = 18, angle = 45, hjust = 1),
    legend.position = "none"
  )

print(mhWinMod_fixEffBar)
```

### Random Effect Slope Distributions

These histograms show the distribution of individual differences in how each predictor affects choice behavior.

```{r mh-random-slopes-ptr, fig.width=4, fig.height=6}
# Post-tournament rating random slopes
# Shows individual variation in sensitivity to learned outcome probabilities
bin_count <- 20
x_min <- min(outProbEstRfx)
x_max <- max(outProbEstRfx)
bin_width <- (x_max - x_min) / bin_count

hist_ptr <- ggplot(outProbEstRfx, aes(x = outProbRfx)) +
  geom_histogram(binwidth = bin_width, fill = "#74ADD1", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.2, x_max*1.3)) +
  labs(
    x = "Post-tournament rating slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)

print(hist_ptr)
```

```{r mh-random-slopes-valence, fig.width=4, fig.height=6}
# Valence random slopes
# Shows individual variation in reward vs. loss sensitivity
bin_count <- 20
x_min <- min(valRfx)
x_max <- max(valRfx)
bin_width <- (x_max - x_min) / bin_count

hist_val <- ggplot(valRfx, aes(x = random_effects_mh$valence1)) +
  geom_histogram(binwidth = bin_width, fill = "#A6D854", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.1, x_max*1.1)) +
  labs(
    x = "Valence slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)

print(hist_val)
```

```{r mh-random-slopes-effort, fig.width=4, fig.height=6}
# Effort level random slopes
# Shows individual variation in effort sensitivity
bin_count <- 20
x_min <- min(effLvlRfx)
x_max <- max(effLvlRfx)
bin_width <- (x_max - x_min) / bin_count

hist_effLvl <- ggplot(effLvlRfx, aes(x = random_effects_mh$effPrp)) +
  geom_histogram(binwidth = bin_width, fill = "#66C2A5", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.3, x_max*1.3)) +
  labs(
    x = "Effort level slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)

print(hist_effLvl)
```

```{r mh-random-slopes-magnitude, fig.width=4, fig.height=6}
# Outcome magnitude random slopes
# Shows individual variation in magnitude sensitivity
bin_count <- 20
x_min <- min(outMagRfx)
x_max <- max(outMagRfx)
bin_width <- (x_max - x_min) / bin_count

hist_outMag <- ggplot(outMagRfx, aes(x = random_effects_mh$normOutMag)) + 
  geom_histogram(binwidth = bin_width, fill = "#FEE08B", color = "black", linewidth = 0.9) +
  scale_x_continuous(limits = c(x_min*1.3, x_max*1.3)) +
  labs(
    x = "Outc. magn. slopes",
    y = "Frequency") +
  theme_minimal(base_size = 16)

print(hist_outMag)
```

---

## SHAPS Interaction Models

Testing whether SHAPS (anhedonia) interacts with different decision variables. Each model tests a specific interaction while controlling for main effects.

```{r shaps-interaction-models}
# Post-tournament rating × SHAPS interaction
# Tests whether anhedonia modulates sensitivity to learned outcome probabilities
intrMod_ptr_SHAPS <- glmer(accepted ~ Age_resc + Gender + normOutMag + effPrp + valence + postTournRat * SHAPS_resc 
                           + (1 + postTournRat + normOutMag + valence + effPrp | prolificID), 
                           data = tbtDf, 
                           family = binomial(link = "logit"), 
                           control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 2e5)))
summary(intrMod_ptr_SHAPS)
```

```{r shaps-valence-interaction}
# Valence × SHAPS interaction
# Tests whether anhedonia differentially affects reward vs. loss processing
intrMod_val_SHAPS <- glmer(accepted ~ Age_resc + Gender + normOutMag + effPrp + postTournRat + valence * SHAPS_resc 
                           + (1 + postTournRat + normOutMag + valence + effPrp | prolificID), 
                           data = tbtDf, 
                           family = binomial(link = "logit"), 
                           control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 2e5)))
summary(intrMod_val_SHAPS)
```

```{r shaps-effort-interaction}
# Effort × SHAPS interaction
# Tests whether anhedonia modulates effort sensitivity
intrMod_eff_SHAPS <- glmer(accepted ~ Age_resc + Gender + normOutMag + postTournRat + effPrp * SHAPS_resc 
                           + (1 + postTournRat + normOutMag + valence + effPrp | prolificID), 
                           data = tbtDf, 
                           family = binomial(link = "logit"), 
                           control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 2e5)))
summary(intrMod_eff_SHAPS)
```

```{r shaps-magnitude-interaction}
# Outcome magnitude × SHAPS interaction
# Tests whether anhedonia modulates magnitude sensitivity
intrMod_mag_SHAPS <- glmer(accepted ~ Age_resc + Gender + effPrp + postTournRat + valence + normOutMag * SHAPS_resc 
                           + (1 + postTournRat + normOutMag + valence + effPrp | prolificID), 
                           data = tbtDf, 
                           family = binomial(link = "logit"), 
                           control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 2e5)))
summary(intrMod_mag_SHAPS)
```

---

## SHAPS × Outcome Probability Interaction

```{r shaps-interaction, fig.width=10, fig.height=6}
if ("SHAPS_resc" %in% names(coef(mh_fits$best_model)$prolificID)) {
  
  # Create a grid of values for prediction
  # This generates combinations of predictor values to visualize the interaction
  newdata <- expand.grid(
    postTournRat = seq(0, 1, by = 0.1),          # Range of post-tournament ratings
    SHAPS_resc = c(-1.5, 0, 1.5),                # Low, mean, and high anhedonia
    Age_resc = 0,                                 # Set to mean (since rescaled)
    Gender = "Woman (including Trans Female/Trans Woman)",  # Most common category
    effPrp = mean(tbtDf$effPrp, na.rm = TRUE),   # Average effort level
    normOutMag = mean(tbtDf$normOutMag, na.rm = TRUE),  # Average outcome magnitude
    valence = levels(tbtDf$valence)[1]           # Use first level of valence
  )
  
  # Predict acceptance probabilities for each combination
  # re.form = NA means we use population-level (fixed) effects only
  newdata$predicted <- predict(mh_fits$best_model, newdata = newdata, type = "response", re.form = NA)
  
  # Convert SHAPS_resc to factor for better labeling in plots
  newdata$SHAPS_resc_factor <- factor(newdata$SHAPS_resc, 
                                      levels = c(-1.5, 0, 1.5),
                                      labels = c("Low", "Mean", "High"))
  
  # Create line plot showing interaction
  # Shows how the effect of post-tournament rating varies by SHAPS level
  plot2 <- ggplot(newdata, aes(x = postTournRat, y = predicted, color = SHAPS_resc_factor)) +
    geom_line(linewidth = 1.5) +
    scale_color_manual(values = c("#ef476f", "#ffd166", "#26547c"),
                       name = "SHAPS Score") +
    labs(x = "Post-tournament Rating",
         y = "Probability of Acceptance") +
    theme_minimal(base_size = 18) +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          legend.title = element_text(face = "bold"))
  
  print(plot2)
  
  # Additional visualization: Showing the difference in slopes more explicitly
  # This bar plot makes it easier to see how anhedonia affects the slope
  # Slope values calculated as: base effect ± interaction effect
  slopes <- data.frame(
    SHAPS_level = c("Low", "Mean", "High"),
    Slope = c(6.02397 - 0.71045, 6.02397, 6.02397 + 0.71045)
  )
  
  # Create bar plot comparing slopes at different SHAPS levels
  plot3 <- ggplot(slopes, aes(x = SHAPS_level, y = Slope, fill = SHAPS_level)) +
    geom_col(width = 0.6) +
    scale_fill_manual(values = c("#26547c", "#ef476f", "#ffd166")) +  
    labs(x = "SHAPS Score",
         y = "Effect of Post-tourn. rating on Accepting Effort") +
    theme_minimal(base_size = 18) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank()) +
    geom_text(aes(label = round(Slope, 2)), vjust = -0.5, size = 4)
  
  print(plot3)
  
  # Combine both plots side by side for comprehensive visualization
  comb_intr_plot <- gridExtra::grid.arrange(plot2, plot3, ncol = 2)
  
} else {
  cat("SHAPS not included in best mental health model\n")
}
```

---

# Descriptive Visualizations

## Learning Trajectories

Shows how participants learned stimulus values over trials.

```{r learning-trajectories, fig.width=10, fig.height=12}
if (file.exists(file.path(data_dir, "allEstsByStim.csv"))) {
  
  # Read the raw estimates data.
  # header = FALSE is important here: this file has no column-name row,
  # it's pure numbers from the first line. Without header = FALSE, read.csv()
  # would wrongly treat that first data row as column headers and silently
  # drop it from the analysis.
  rawEstData <- read.csv(file.path(data_dir, "allEstsByStim.csv"), header = FALSE)
  
  # Convert raw matrix to tibble
  df <- as_tibble(rawEstData)
  
  # Rename columns for clarity
  colnames(df) <- c("stimulus", paste0("t", 1:(ncol(df)-1)))
  
  # Pivot to long format: one row per participant × trial
  long_df <- df %>%
    pivot_longer(
      cols = starts_with("t"),
      names_to = "trial",
      names_prefix = "t",
      names_transform = list(trial = as.integer),
      values_to = "estimate"
    )
  
  # Calculate mean and SE per stimulus × trial
  summary_df <- long_df %>%
    group_by(stimulus, trial) %>%
    dplyr::summarize(
      mean_est = mean(estimate, na.rm = TRUE),
      se_est = sd(estimate, na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )
  
  # Plot learning trajectories
  p <- ggplot(summary_df, aes(x = trial, y = mean_est, color = factor(stimulus), group = factor(stimulus))) +
    geom_line(linewidth = 0.5) +
    geom_point(size = 1.2) +
    geom_ribbon(aes(ymin = mean_est - se_est, ymax = mean_est + se_est, fill = factor(stimulus)), 
                alpha = 0.4, color = NA) +
    geom_hline(data = data.frame(
      stimulus = 1:8,
      h_line = ifelse(c(1:8) %in% c(1, 3, 5, 7), 0.25, 0.75)), 
      aes(yintercept = h_line), linetype = "dashed", color = "black") +
    scale_x_continuous(breaks = seq(1, 16, by = 3)) +
    facet_wrap(~ stimulus, ncol = 2, labeller = labeller(stimulus = function(x) paste("Stimulus", x))) +
    ylim(0, 1) +
    labs(
      x = "Trial",
      y = "Average Rating",
      color = "Stimulus",
      fill = "Stimulus"
    ) +
    theme_minimal(base_size = 20) +
    theme(
      legend.position = "bottom",
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "lightgray", color = NA),
      strip.text = element_text(face = "bold")
    )
  
  print(p)
  
} else {
  cat("Learning trajectory data (data/allEstsByStim.csv) not found\n")
}
```

## Post-Tournament Estimation Error

Shows the error in participants' post-tournament outcome probability estimates. Each point represents one participant's average estimation error. The yellow line at 0 indicates perfect accuracy.

```{r post-tournament-error-scatter, fig.width=10, fig.height=6}
# Create scatter plot of post-tournament estimation errors
# Each point represents one participant's average error across all stimuli
# Positive values = overestimation, negative values = underestimation
postTournErrScat <- ggplot(summDf, aes(x = prolificID, y = postTournErr)) +
  geom_point(color = "black") +
  geom_hline(yintercept = 0, color = "#fed700", linetype = "solid", linewidth = 1.5) +
  labs(
    x = "Participants",
    y = "Error in post-tournament estimates") +
  theme_pubclean(base_size = 18) +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 18),
        axis.ticks.x = element_blank(),
        axis.text.x.bottom = element_blank()) +
  ylim(-0.4, 0.4)

print(postTournErrScat)
```

## Final Learning Estimates

Shows how well participants learned each stimulus by the end of training.

```{r final-estimates, fig.width=10, fig.height=6}
# Reorder variables: odd numbers (1,3,5,7) then even numbers (2,4,6,8)
variables_to_mean <- c("finLearningEst_1", "finLearningEst_3", "finLearningEst_5", "finLearningEst_7",
                       "finLearningEst_2", "finLearningEst_4", "finLearningEst_6", "finLearningEst_8")

# Convert to long format with reordered variables
df_long <- summDf %>%
  tidyr::pivot_longer(cols = all_of(variables_to_mean), 
                      names_to = "Variable", 
                      values_to = "Value") %>%
  mutate(
    Variable = factor(Variable, levels = variables_to_mean),
    Valence = case_when(
      Variable %in% c("finLearningEst_1", "finLearningEst_2", "finLearningEst_3", "finLearningEst_4") ~ "Reward",
      Variable %in% c("finLearningEst_5", "finLearningEst_6", "finLearningEst_7", "finLearningEst_8") ~ "Loss"
    )
  )

# Custom x-axis labels
custom_labels <- c("Alien 1", "Alien 3", "Alien 5", "Alien 7",
                   "Alien 2", "Alien 4", "Alien 6", "Alien 8")

# Bright red and green colors
condition_colors <- c("Loss" = "#FF1744", "Reward" = "#1AC71A")

# Create publication-ready boxplot with scatter points
p <- ggplot(df_long, aes(x = Variable, y = Value, fill = Valence, color = Valence)) +
  geom_boxplot(alpha = 0.5) +
  geom_point(position = position_jitter(width = 0.1), alpha = 0.8, size = 1) +
  scale_fill_manual(values = condition_colors) +
  scale_color_manual(values = condition_colors) +
  scale_x_discrete(labels = custom_labels) +
  labs(x = "Stimulus", y = "Final learning estimate") +
  theme_minimal(base_size = 18) +
  ylim(0,1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
    axis.text.y = element_text(size = 16),
    axis.title = element_text(size = 18),
    plot.title = element_text(size = 14, hjust = 0.5),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )
print(p)
```

## Questionnaire Distributions

```{r questionnaire-distributions, fig.width=8, fig.height=6}
# Prepare data for SHAPS and FAS
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", measure.vars = c("SHAPS","FAS"))

# Create raincloud plot
shapsFasRain <- ggplot(summ_long, aes(x = variable, y = value, fill = as.factor(variable))) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) +
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .2,
    position = position_jitter(
      seed = 1, width = .1
    ) 
  ) + 
  geom_boxplot(
    width = .22, 
    outlier.shape = NA,
    aes(fill = as.factor(variable))) +
  coord_cartesian(xlim = c(1.2, 2.4), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  xlim(0,50) +
  labs(
    x = "",
    y = "Questionnaire score") +
  theme_minimal(base_size = 18) +
  theme(panel.background = element_rect(fill = "white", colour = "white")) +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("SHAPS" = "SHAPS", "FAS" = "FAS")) +
  scale_fill_manual(values = c("SHAPS" = "#30c5d2","FAS" = "#E4C2C6")) +
  scale_color_manual(values = c("SHAPS" = "#30c5d2","FAS" = "#E4C2C6"))

print(shapsFasRain)
```

```{r transdiagnostic-distributions, fig.width=8, fig.height=6}
# Prepare data for transdiagnostic factors
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", measure.vars = c("AD","SW","Compul"))

# Create raincloud plot
transxRain <- ggplot(summ_long, aes(x = variable, y = value, fill = as.factor(variable))) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) +
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .2,
    position = position_jitter(
      seed = 1, width = .1
    ) 
  ) + 
  geom_boxplot(
    width = .22, 
    outlier.shape = NA,
    aes(fill = as.factor(variable))) +
  coord_cartesian(xlim = c(1.2, 3.4), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    x = "",
    y = "Factor loading") +
  theme_minimal(base_size = 18) +
  theme(panel.background = element_rect(fill = "white", colour = "white")) +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("AD" = "AD", "SW" = "SW", "Compul" = "Compul")) +
  scale_fill_manual(values = c("AD" = "#84ffc9","SW" = "#aab2ff", "Compul" = "#eca0ff")) +
  scale_color_manual(values = c("AD" = "#84ffc9","SW" = "#aab2ff", "Compul" = "#eca0ff"))

print(transxRain)
```

---

# Statistical Tests

## Partial Correlations

Examining relationships between anhedonia (SHAPS) and effort acceptance, controlling for age and gender.

```{r partial-correlations}
# Define control variables
control_vars <- cbind(as.numeric(summDf$Age), as.factor(summDf$Gender))

# Initialize results dataframe
pcor_results <- data.frame(
  comparison = character(),
  estimate = numeric(),
  p_value = numeric(),
  statistic = numeric(),
  stringsAsFactors = FALSE
)

# Low effort
if ("percAcceptByEffort_1" %in% names(summDf)) {
  pacc_low <- asin(sqrt(summDf$percAcceptByEffort_1 / 100))
  pcor_low <- pcor.test(summDf$SHAPS_resc, pacc_low, control_vars, method = "kendall")
  pcor_results <- rbind(pcor_results, data.frame(
    comparison = "SHAPS × Low effort acceptance",
    estimate = pcor_low$estimate,
    p_value = pcor_low$p.value,
    statistic = pcor_low$statistic
  ))
}

# Medium effort
if ("percAcceptByEffort_2" %in% names(summDf)) {
  pacc_med <- asin(sqrt(summDf$percAcceptByEffort_2 / 100))
  pcor_med <- pcor.test(summDf$SHAPS_resc, pacc_med, control_vars, method = "kendall")
  pcor_results <- rbind(pcor_results, data.frame(
    comparison = "SHAPS × Medium effort acceptance",
    estimate = pcor_med$estimate,
    p_value = pcor_med$p.value,
    statistic = pcor_med$statistic
  ))
}

# High effort
if ("percAcceptByEffort_3" %in% names(summDf)) {
  pacc_high <- asin(sqrt(summDf$percAcceptByEffort_3 / 100))
  pcor_high <- pcor.test(summDf$SHAPS_resc, pacc_high, control_vars, method = "kendall")
  pcor_results <- rbind(pcor_results, data.frame(
    comparison = "SHAPS × High effort acceptance",
    estimate = pcor_high$estimate,
    p_value = pcor_high$p.value,
    statistic = pcor_high$statistic
  ))
}

# Overall acceptance
if ("percAccept" %in% names(summDf)) {
  pcor_overall <- pcor.test(summDf$SHAPS_resc, summDf$percAccept, control_vars, method = "kendall")
  pcor_results <- rbind(pcor_results, data.frame(
    comparison = "SHAPS × Overall acceptance",
    estimate = pcor_overall$estimate,
    p_value = pcor_overall$p.value,
    statistic = pcor_overall$statistic
  ))
}

print(pcor_results)

```

## Supplementary Figure 6 - SHAPS Effects on Effort Acceptance

### SHAPS × Effort Level Interaction

Shows how anhedonia (SHAPS) moderates the effect of effort level on acceptance probability. Three lines represent low (-1.5 SD), mean (0), and high (3 SD) SHAPS scores.

```{r supp-fig-6a-shaps-effort, fig.width=10, fig.height=6}
# Generate marginal effects predictions for SHAPS × effort interaction
# Shows predicted probability of accepting effort at different SHAPS levels
# type = "fe" means fixed effects only (population-level predictions)

mhmod <- mh_fits$best_model
mhMod_SHAPS_ptr_plot <- ggpredict(mhmod, 
                                  terms = c("effPrp [all]",           # Range of effort levels
                                           "SHAPS_resc [-1.5, 0, 3]"), # Low, mean, high SHAPS
                                  type = "fixed") %>%
  plot() +
  # Custom color scheme for SHAPS levels
  scale_fill_manual(values = c("#F9AED3",      # Low SHAPS (pink)
                               "#8AC926",      # Mean SHAPS (green)
                               "#4A90E0")) +   # High SHAPS (blue)
  scale_colour_manual(values = c("#F9AED3", "#8AC926", "#4A90E0")) +
  labs(
    title = "",
    x = "Effort level",
    y = "p(accept)",
    colour = "SHAPS"
  ) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "right")

print(mhMod_SHAPS_ptr_plot)
```

### SHAPS Main Effect on Acceptance Probability

Shows the main effect of anhedonia on overall effort acceptance probability, controlling for all other variables in the model.

```{r supp-fig-6b-shaps-main-effect, fig.width=10, fig.height=6}
# Generate marginal effects for SHAPS main effect
# Shows how acceptance probability changes across the full range of SHAPS scores
marginal_effects <- ggpredict(mh_fits$best_model, terms = "SHAPS_resc [all]")

# Create line plot with confidence ribbon
# Pink color emphasizes the anhedonia measure
mhWinMod_SHAPS_pAccept_plot <- ggplot(marginal_effects, aes(x = x, y = predicted)) +
  # Main effect line
  geom_line(color = "#F9AED3", linewidth = 2) +
  # 95% confidence interval ribbon
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.2,           # Transparency for ribbon
              color = "#F9AED3",     # Ribbon border color
              fill = "#F9AED3") +    # Ribbon fill color
  labs(
    x = "SHAPS score (z-scored)",
    y = "p(accept)"
  ) +
  theme_minimal(base_size = 18)

print(mhWinMod_SHAPS_pAccept_plot)
```

---

# Summary

## Sample Characteristics

```{r summary-stats}
summary_stats <- data.frame(
  Measure = c("Total participants", "Total trials", "Mean age (SD)", 
              "Gender (% female)", "Mean SHAPS (SD)", "Mean FAS (SD)"),
  Value = c(
    length(unique(tbtDf$prolificID)),
    nrow(tbtDf),
    sprintf("%.2f (%.2f)", mean(summDf$Age, na.rm = TRUE), sd(summDf$Age, na.rm = TRUE)),
    sprintf("%.2f%%", mean(summDf$Gender == "Woman (including Trans Female/Trans Woman)", na.rm = TRUE) * 100),
    sprintf("%.2f (%.2f)", mean(summDf$SHAPS, na.rm = TRUE), sd(summDf$SHAPS, na.rm = TRUE)),
    sprintf("%.2f (%.2f)", mean(summDf$FAS, na.rm = TRUE), sd(summDf$FAS, na.rm = TRUE))
  )
)

print(summary_stats)
```

# Session Information

```{r session-info}
sessionInfo()
```
