---
title: "LxM Supplementary Figures"
author: "Calum Guinea"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)
```

# Overview

This notebook generates all publication-ready figures from the LxM (Learning × Motivation) manuscript. Each figure is clearly labeled and organized by manuscript section (main figures and supplementary figures).

**To use this notebook:**

1. Update the working directory in the "Set Working Directory" chunk
2. Ensure your data files and preprocessing script are in that directory
3. Run all chunks sequentially (or use "Run All")

---

# Setup

## Set Working Directory

```{r set-directory}
# Change this to your data folder location
setwd('C:/Users/calum/desktop')
```

## Install Packages (if needed)

Run this chunk once if you need to install packages:

```{r install-packages, eval=FALSE}
install.packages(c("dplyr", "tidyr", "ggplot2", "ggdist", "gghalves", 
                   "ggbeeswarm", "cowplot", "ggpubr", "RColorBrewer", 
                   "reshape2", "colorspace", "ggsci", "corrplot"))
```

## Load Required Packages

```{r load-packages}
library(dplyr)       # Data manipulation
library(tidyr)       # Data tidying
library(ggplot2)     # Plotting
library(ggdist)      # Distribution visualizations
library(gghalves)    # Half-violin plots
library(ggbeeswarm)  # Beeswarm plots
library(cowplot)     # Publication-ready themes
library(ggpubr)      # Publication-ready plots
library(RColorBrewer)# Color palettes
library(reshape2)    # Data reshaping
library(colorspace)  # Color utilities
library(ggsci)       # Scientific journal color palettes
library(corrplot)    # Correlation matrix plots
```

---

# Data Loading and Preprocessing

## Define Preprocessing Function

This function loads the data files and prepares them for analysis by:

- Converting valence to binary (0 = loss, 1 = reward)
- Scaling continuous predictors (mean-centering and standardizing)
- Optionally excluding participants with poor tournament performance

```{r preprocessing-function}
load_and_preprocess_data <- function(exclude_low_performance = TRUE) {
  
  # Read data files
  tbtDf <- read.csv("choice_trialByTrial.csv")
  summDf <- read.csv("summaryFull.csv")
  
  # Rename ID column for consistency
  summDf <- summDf %>% rename(prolificID = ParticipantId)
  
  # Convert valence to binary factor (0 = loss, 1 = reward)
  tbtDf <- tbtDf %>% mutate(valence = ifelse(valence == -1, 0, 1))
  tbtDf$valence <- as.factor(tbtDf$valence)
  
  # Scale continuous predictors (mean-centered and standardized)
  # Learning fidelity measures
  tbtDf$postLearnRat_resc <- scale(tbtDf$postLearnRat)
  tbtDf$postTournRat_resc <- scale(tbtDf$postTournRat)
  tbtDf$postEffortRat_resc <- scale(tbtDf$postEffortRat)
  tbtDf$meanEstByStim_resc <- scale(tbtDf$meanEstByStim)
  tbtDf$finLearningEst_resc <- scale(tbtDf$finLearningEst)
  
  # Decision variables
  tbtDf$effDiscEV <- scale(tbtDf$effDiscEV)
  tbtDf$alienEV <- scale(tbtDf$alienEV)
  
  # Demographics
  tbtDf$Age_resc <- scale(tbtDf$Age)
  tbtDf$Sex <- as.factor(tbtDf$Sex)
  tbtDf$Gender <- as.factor(tbtDf$Gender)
  
  # Questionnaire scores
  tbtDf$SHAPS_resc <- scale(tbtDf$SHAPS)
  summDf$SHAPS_resc <- scale(summDf$SHAPS)
  tbtDf$AD_resc <- scale(tbtDf$AD)
  tbtDf$SW_resc <- scale(tbtDf$SW)
  tbtDf$Compul_resc <- scale(tbtDf$Compul)
  tbtDf$FAS_resc <- scale(tbtDf$FAS)
  tbtDf$AES_resc <- scale(tbtDf$AES)
  tbtDf$STAI_resc <- scale(tbtDf$STAI)
  
  summDf$FAS_resc <- scale(summDf$FAS)
  summDf$AES_resc <- scale(summDf$AES)
  summDf$STAI_resc <- scale(summDf$STAI)
  
  # Exclude participants with poor tournament performance if requested
  if (exclude_low_performance) {
    threshold <- 50  # Above chance performance
    tbtDf <- tbtDf[tbtDf$ovrperc > threshold, ]
    summDf <- summDf[summDf$ovrperc > threshold, ]
  }
  
  # Create expected value difference variable
  tbtDf <- tbtDf %>%
    mutate(evDiff = ifelse(
      outMag < 0,
      (abs(outMag) * outProb) - (abs(outMag) * 1),
      (abs(outMag) * outProb) - (abs(outMag) * 0)
    ))
  tbtDf$evDiff <- abs(tbtDf$evDiff)
  
  return(list(tbtDf = tbtDf, summDf = summDf))
}
```

## Load Data

```{r load-data}
data <- load_and_preprocess_data(exclude_low_performance = TRUE)
tbtDf <- data$tbtDf
summDf <- data$summDf
```

---

# Supplementary Figures

## Supplementary Figure 1 - Maximum Voluntary Contraction (MVC)

Shows each participant's maximum key presses per second. The horizontal line at 19 indicates the minimum threshold used to exclude participants with insufficient effort capacity.

```{r supp-fig-1-mvc, fig.width=10, fig.height=6}
# Create bar plot showing MVC for each participant
# MVC = Maximum Voluntary Contraction (max key presses per second)
mvcBar <- ggplot(summDf, aes(x = prolificID, y = mvc, fill = "#ff6201")) +
  # Create bar for each participant
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  # Add horizontal line showing minimum threshold (19 presses)
  geom_hline(yintercept = 19, color = "black", linetype = "solid", linewidth = 1.25) +
  scale_fill_manual(values = "#ff6201") +
  labs(
    x = "Participants",
    y = "Maximum key presses") +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),              # Hide participant IDs
    panel.background = element_rect(fill = "white", colour = "white"),
    axis.ticks.x = element_blank(),
    axis.text.x.bottom = element_blank()) +
  ylim(0, 100)  # Set y-axis limits

print(mvcBar)
```

## Supplementary Figure 2 - Mean Post-Task Estimates

Raincloud plots showing distribution of mean post-task probability estimates across three timepoints: post-learning, post-tournament, and post-effort.

```{r supp-fig-2-post-task-estimates, fig.width=10, fig.height=6}
# Prepare data for plotting
# Select three post-task estimate measures and convert to long format
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", 
                             measure.vars = c("postLearnMean_1", 
                                             "postTournMean_1", 
                                             "postEffortMean_1"))

# Create raincloud plot (combination of violin, boxplot, and points)
# This shows the full distribution at each timepoint
meanPostEstsRain <- ggplot(summ_long, aes(x = variable, y = value, 
                                          fill = as.factor(variable))) + 
  # Add half-violin plot showing distribution density
  ggdist::stat_halfeye(
    adjust = .5,           # Bandwidth adjustment for smoothing
    width = .6,            # Width of the violin
    .width = 0,            # Don't show interval
    justification = -.3,   # Position relative to box
    point_colour = NA) +   # No point in violin
  # Add boxplot showing median and quartiles
  geom_boxplot(
    width = .25,           # Width of boxplot
    outlier.shape = NA,    # Hide outliers (shown as points instead)
    aes(fill = as.factor(variable))) +
  # Add individual data points with jitter
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .2,            # Transparency to show overlapping points
    position = position_jitter(seed = 1, width = .1)) + 
  # Adjust plot area to accommodate violin
  coord_cartesian(xlim = c(1.2, 3.2), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  labs(y = "Mean post-task estimates (as proportions)") +
  theme_minimal(base_size = 18) +
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    legend.position = "none",
    axis.title.x = element_blank()) +
  # Custom x-axis labels
  scale_x_discrete(labels = c("postLearnMean_1" = "Post-learning", 
                              "postTournMean_1" = "Post-tournament",
                              "postEffortMean_1" = "Post-effort")) +
  # Custom color scheme for each timepoint
  scale_fill_manual(values = c("postLearnMean_1" = "#6dd3ce",
                               "postTournMean_1" = "#c8e9a0",
                               "postEffortMean_1" = "#f7a278")) +
  scale_color_manual(values = c("postLearnMean_1" = "#6dd3ce",
                                "postTournMean_1" = "#c8e9a0",
                                "postEffortMean_1" = "#f7a278")) +
  # Set y-axis to show probabilities from 0 to 1
  scale_y_continuous(
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    labels = c("0", "0.25", "0.5", "0.75", "1"))

print(meanPostEstsRain)
```

## Supplementary Figure 3 - Tournament Performance

Raincloud plots showing accuracy across different tournament trial types: overall, reward-only, loss-only, approach, and avoid trials.

```{r supp-fig-3-tournament-performance, fig.width=10, fig.height=6}
# Prepare tournament performance data
# Five measures of tournament accuracy across different trial types
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", 
                             measure.vars = c("ovrperc",      # Overall accuracy
                                             "rewOvrPerc",    # Reward trials only
                                             "lossOvrPerc",   # Loss trials only
                                             "appPerc",       # Approach trials
                                             "avPerc"))       # Avoid trials

# Create raincloud plot
tournAccRain <- ggplot(summ_long, aes(x = variable, y = value, 
                                      fill = as.factor(variable))) + 
  # Half-violin showing distribution
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) +
  # Individual data points
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .5,
    position = position_jitter(seed = 1, width = .1)) + 
  # Boxplot overlay
  geom_boxplot(
    width = .25, 
    outlier.shape = NA,
    aes(fill = as.factor(variable))) +
  coord_cartesian(xlim = c(1.2, 5.4), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    y = "Tournament performance (as %)",
    x = "Accuracy") +
  theme_minimal(base_size = 18) +
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
    axis.text.y = element_text(size = 18)) +
  # Custom colors for each trial type
  scale_fill_manual(values = c("ovrperc" = "#fb8fa4",
                               "rewOvrPerc" = "#7dbedf",
                               "lossOvrPerc" = "#f7cd4d",
                               "appPerc" = "#91c579",
                               "avPerc" = "#d2aecc")) +
  scale_color_manual(values = c("ovrperc" = "#fb8fa4",
                                "rewOvrPerc" = "#7dbedf",
                                "lossOvrPerc" = "#f7cd4d",
                                "appPerc" = "#91c579",
                                "avPerc" = "#d2aecc")) +
  # Custom x-axis labels
  scale_x_discrete(labels = c("ovrperc" = "Overall", 
                              "rewOvrPerc" = "Reward-only", 
                              "lossOvrPerc" = "Loss-only", 
                              "appPerc" = "Approach", 
                              "avPerc" = "Avoid"))

print(tournAccRain)
```

## Supplementary Figure 4 - Correlation Matrix for Learning Measures

Correlation matrix showing relationships among five individualized learning measures. This helps validate that different learning measures capture similar underlying constructs.

```{r supp-fig-4-correlation-matrix, fig.width=9, fig.height=6}
# Compute the mean for each learning measure, grouped by participant
# This creates one value per participant per measure
rat_means <- tbtDf %>%
  group_by(prolificID) %>%
  summarise(
    plrMean = mean(postLearnRat, na.rm = TRUE),      # Post-learning rating
    ptrMean = mean(postTournRat, na.rm = TRUE),      # Post-tournament rating
    perMean = mean(postEffortRat, na.rm = TRUE),     # Post-effort rating
    fleMean = mean(as.numeric(finLearningEst), na.rm = TRUE),  # Final learning estimate
    mebsMean = mean(meanEstByStim, na.rm = TRUE)     # Mean trial-by-trial estimate
  )

# Convert to dataframe
rat_means <- as.data.frame(rat_means)

# Select only the learning measure columns (exclude prolificID)
corr_data <- rat_means[, c("plrMean", "ptrMean", "perMean", "fleMean", "mebsMean")]

# Compute Pearson correlation matrix
# use = "pairwise.complete.obs" handles missing data
cor_matrix <- cor(corr_data, use = "pairwise.complete.obs")

# Create readable labels for the correlation matrix
custom_labels <- c("Post Learning Mean", 
                   "Post Tourn. Mean", 
                   "Post Effort Mean", 
                   "Final Learning Estimate", 
                   "Mean tbt estimate")

# Apply custom labels to matrix
rownames(cor_matrix) <- custom_labels
colnames(cor_matrix) <- custom_labels

# Function to compute p-values for correlations
# Tests significance of each pairwise correlation
cor_pvalues <- function(mat) {
  n <- ncol(mat)
  p_mat <- matrix(NA, n, n)
  rownames(p_mat) <- colnames(mat)
  colnames(p_mat) <- colnames(mat)
  
  # Test each unique pair
  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      test <- cor.test(mat[, i], mat[, j], 
                      use = "pairwise.complete.obs", 
                      method = "pearson")
      p_mat[i, j] <- test$p.value
      p_mat[j, i] <- test$p.value  # Matrix is symmetric
    }
  }
  
  diag(p_mat) <- 0  # Set diagonal to 0
  return(p_mat)
}

# Compute p-values for significance testing
p_matrix <- cor_pvalues(corr_data)

# Generate the correlation matrix heatmap
# Shows lower triangle only (matrix is symmetric)
# Color gradient: blue (negative) → white (zero) → red (positive)
corrplot(cor_matrix, 
         method = "color",                      # Use color squares
         type = "lower",                        # Show lower triangle only
         col = colorRampPalette(c("blue", "white", "red"))(200),  # Color gradient
         tl.col = "black",                      # Text label color
         tl.cex = 0.8,                         # Text label size
         tl.srt = 45,                          # Text label rotation
         addCoef.col = "black")                # Add correlation coefficients
```

## Supplementary Figure 5 - Questionnaire Distributions

Raincloud plots showing distributions of questionnaire scores: SHAPS (anhedonia), FAS (fatigue), and transdiagnostic factors (Anxious-Depression, Social Withdrawal, Compulsivity).

### SHAPS and FAS

```{r supp-fig-5a-shaps-fas, fig.width=8, fig.height=6}
# Prepare SHAPS and FAS data
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", 
                             measure.vars = c("SHAPS", "FAS"))

# Create raincloud plot for SHAPS and FAS
shapsFasRain <- ggplot(summ_long, aes(x = variable, y = value, 
                                      fill = as.factor(variable))) + 
  # Half-violin distribution
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) +
  # Individual points with jitter
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .2,
    position = position_jitter(seed = 1, width = .1)) + 
  # Boxplot overlay
  geom_boxplot(
    width = .22, 
    outlier.shape = NA,
    aes(fill = as.factor(variable))) +
  coord_cartesian(xlim = c(1.2, 2.4), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    x = "",
    y = "Questionnaire score") +
  theme_minimal(base_size = 18) +
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    legend.position = "none") +
  scale_x_discrete(labels = c("SHAPS" = "SHAPS", "FAS" = "FAS")) +
  # Custom colors: cyan for SHAPS, pink for FAS
  scale_fill_manual(values = c("SHAPS" = "#30c5d2", "FAS" = "#E4C2C6")) +
  scale_color_manual(values = c("SHAPS" = "#30c5d2", "FAS" = "#E4C2C6"))

print(shapsFasRain)
```

### Transdiagnostic Factors

```{r supp-fig-5b-transdiagnostic, fig.width=8, fig.height=6}
# Prepare transdiagnostic factor data
# AD = Anxious-Depression
# SW = Social Withdrawal
# Compul = Compulsivity
summ_long <- reshape2::melt(summDf, id.vars = "prolificID", 
                             measure.vars = c("AD", "SW", "Compul"))

# Create raincloud plot for transdiagnostic factors
transxRain <- ggplot(summ_long, aes(x = variable, y = value, 
                                    fill = as.factor(variable))) + 
  # Half-violin distribution
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) +
  # Individual points with jitter
  geom_point(
    inherit.aes = T,
    aes(color = as.factor(variable)),
    size = 1.5,
    alpha = .2,
    position = position_jitter(seed = 1, width = .1)) + 
  # Boxplot overlay
  geom_boxplot(
    width = .22, 
    outlier.shape = NA,
    aes(fill = as.factor(variable))) +
  coord_cartesian(xlim = c(1.2, 3.4), clip = "off") +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    x = "",
    y = "Factor loading") +
  theme_minimal(base_size = 18) +
  theme(
    panel.background = element_rect(fill = "white", colour = "white"),
    legend.position = "none") +
  scale_x_discrete(labels = c("AD" = "AD", "SW" = "SW", "Compul" = "Compul")) +
  # Custom colors for each factor
  scale_fill_manual(values = c("AD" = "#84ffc9",      # Green
                               "SW" = "#aab2ff",       # Blue
                               "Compul" = "#eca0ff")) + # Purple
  scale_color_manual(values = c("AD" = "#84ffc9", 
                                "SW" = "#aab2ff", 
                                "Compul" = "#eca0ff"))

print(transxRain)
```

---

# Session Information

```{r session-info}
sessionInfo()
```
